\documentclass[12pt]{article}
\pagestyle{empty}

\usepackage{fullpage,amssymb,amsmath,amsthm}
\usepackage{cyrillic}
\def\sub{{\rm sub}}
\def\pref{{\rm pref}}
\def\suff{{\rm suff}}
\def\Pref{{\rm Pref}}
\def\shuff{{\rm shuff}}


\thispagestyle{empty}

\begin{document}
\begin{center}
\large\bf University of Waterloo\\
CS~462 --- Formal Languages and Parsing\\
Winter 2013\\
Problem Set 3\\
\end{center}

\bigskip

\noindent{\it Distributed Tuesday, January 22 2013.}

\smallskip

\noindent{\it Due Tuesday, January 29 2013, in class.}

\bigskip\bigskip

All answers should be accompanied by proofs.  

\begin{enumerate}

\item{} [10 marks]  
Must every primitive word of length $> 1$ have at least two 
unbordered conjugates?  Prove or disprove.


\medskip

Yes they do.
\begin{proof}[Proof by Contradiction]$\vDash$
\begin{itemize}
\item[•]Let $w \in \sum^{*}$  and  $\alpha \in \sum$ s.t. $\neg \exists v^{i} = w$, $\mid x \mid >0, \mid x \mid_{\alpha} >0$
\item[•]
\item[•]shuffle($\cdots$, w) := $\bigcup_{W \in shuffle(\cdots)}$shuffle(W, w)
\end{itemize}
\end{proof}

\pagebreak

\item{} [10 marks]  
Consider the equation in words $x \sha y = z^2$, where $\sha$ is the
perfect shuffle.  Describe
all solutions to this equation.  Hint:  there are separate cases
depending on whether $|x|, |y|$ are both even or both odd.

\medskip

(By ``describe all solutions'' we mean come up with a characterization
of the solutions that is somewhat analogous to the descriptions in
Theorems 2.3.2 and 2.3.3 of the course text, giving necessary and
sufficient conditions for $x, y, z$.)

\medskip

if $|x| = |y| = 2 * i$$\vDash$

\begin{itemize}
\item[•]Let x = xa xb and y = ya yb where $|xa| = |xb| = |ya| = |yb| = i$
\item[•]$x \sha y$ = $xa \sha ya$  $xb \sha yb = z z$
\item[•]$xa \sha ya = z = xb \sha yb$
\item[•]xa = xb and ya = yb
\end{itemize}

if $|x| = |y| = 2 * i + 1$$\vDash$
\begin{itemize}
\item[•]Let x = xa xc xb and y = ya yc yb where $|xa| = |xb| = |ya| = |yb| = i and |xc| = |yc| = 1$
\item[•]$x \sha y$ = $xa \sha ya$ xc yc $xb \sha yb$ = z z
\item[•]xa xc = yc yb and ya = xb
\end{itemize}

\medskip

\pagebreak

\item{} [10 marks]  
Call a word $w$ {\it uneven} if every nonempty subword
has the property that at least one letter appears an odd number of
times.  For example, ${\tt abac}$ is uneven.   

\begin{itemize} 

\item[(a)]  [5 marks] Show that if $w$ is an uneven word over an
alphabet with $k$ letters, then $|w| < 2^k$.

\item[(b)] [5 marks]  Prove that the bound in (a) is sharp, by
exhibiting an uneven word of length $2^k -1 $ over every alphabet of
size $k \geq 1$.

\end{itemize}

For (a)
\begin{itemize}

\item[•]Given W $\in \sum^{*}$ and $\mid \sum \mid$ = k
\item[•]Assume $\mid W \mid \ge 2^{k}$

\item[•]Let prefix(w, j) = w' where w' w'' = w and $\mid w' \mid$ = j
\item[•]Let $\bigcup_{i \in Z_{\mid \sum \mid}}\{\alpha_{i}\} = \sum$
\item[•]Let par(w) = v where $v_{i}$ = $\mid w \mid_{\alpha_{i}}$ mod 2
\item[•]Let V = $\{par(prefix(w, j)) \mid j \in Z_{\mid W \mid+1} \setminus \{0\}\}$

if $\exists$ i where $V_{i}$ = par(prefix(W, j)) = 0 then prefix(W, j) is an even word

otherwise $\exists$ i, j where $m \lneq n$, $V_{i}$ = par(prefix(W, m)) = $V_{j}$ = par(prefix(W, n)) because there $2^{k}-1$ unique non-zero vector in the space and $\mid V \mid = \mid W \mid \ge 2^{k}$
\item[•]Let prefix(W, m) z = prefix(W, n)
\item[•]Then par(prefix(W, m)) par(z) = par(prefix(W, n))
\item[•]Then par(z) = par(prefix(W, n)) - par(prefix(W, m)) = 0
\item[•]Then z is an even word

In both cases, W end up contains an even word. Thus $\mid W \mid \not\le 2^{k}$.

\end{itemize}

For (b)
\begin{itemize}
\item[•]Given $\mid \sum \mid$ = k
\item[•]Let $\bigcup_{i \in Z_{\mid \sum \mid}}\{\alpha_{i}\} = \sum$
\item[•]Let $w_{0} = \alpha_{0}$
\item[•]Let $w_{n+1}$ = $w_{n} \sha \alpha_{n+1}^{\mid w_{n} \mid}$ $\alpha_{n+1}$ for $n \ge 0$
\end{itemize}

By observation, $\mid w_{0} \mid = 2^{1} - 1$ and $w_{0}$ is uneven

Given $\mid w_{n} \mid = 2^{n} - 1$ and $w_{n}$ is uneven
\begin{itemize}
\item[•]$\mid w_{n+1} \mid = \mid w_{n+1} \mid * 2 + 1 = 2^{n+1} - 1$
\item[•]Let w be a subword of $w_{n+1}$
\end{itemize}
if $\mid w \mid$ = 1, then w is uneven
otherwise $\mid w \mid \ge 2$, then $w \in perm(w'\alpha_{n+1}^{i})$ where w' is a subword of $w_{n}$
\begin{itemize}
\item[•]Then par(w) = par(w') + $par(\alpha_{n+1}^{i})$
\item[•]Then $par(w') \not= 0$ because w' is uneven
\item[•]Then $par(w') + par(\alpha_{n+1}^{i}) \not= 0$ because $\mid w \mid_{\alpha_{n+1}} = 0$
\item[•]Then w is uneven because $par(w) \not= 0$
In both cases, w is uneven. Therefore $w_{n+1}$ is uneven

By induction, $\mid w_{k} \mid = 2^{k} - 1$ and $w_{k}$ is uneven.

\end{itemize}

\end{enumerate}

\end{document}
